---
title: "ADS2 Group Exercise ICA"
author: "Group 6"
date: "2024-04-02"
output:
  pdf_document:
    fig_height: 4.2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F)
setwd("E:\\A大学\\大二下\\ADS2\\ADS2_ICA")
```

# Part 0 : Clean and tidy the dataset

```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(tidyr)
library(knitr)

substance_use = read.csv("substance_use.csv")
```

**Any missing values?**
  
```{r}
anyNA(substance_use)
sum(substance_use[ , ] == "")
```

There is no missing values in the dataset.

**Any duplicates?**
  
```{r}
which(duplicated(substance_use))
```

There is no duplicate in the dataset.

**Typos, naming schemes and data types?**
  
```{r}
str(substance_use)
```

The location, sex and age variables can be factors.

```{r}
substance_use$location = as.factor(substance_use$location)
substance_use$sex = as.factor(substance_use$sex)
substance_use$age = as.factor(substance_use$age)
```

**Any outliers or strange patterns?**
  
```{r}
ggplot(data = substance_use,
       aes(x = cause, y = val, color = interaction(cause, measure))) +
  geom_boxplot(outlier.size = 0.1) +
  facet_wrap(~location, scales = "free_y") +
  labs(x = "Cause", y = "Value", color = "") +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.key.size = unit(10, "pt"),
        legend.text = element_text(size = 12),
        strip.text = element_text(hjust = 0),
        axis.text.x = element_blank())

which(substance_use$val < 0)
which(substance_use$val > 1)
```

Figure 0: The boxplot of the death rate and prevalence of the alcohol and opioid use disorders in different locations.

The range of the value is normal. From Figure 0, You can see some outliers. However, if you draw the plots for each location, age and sex, you will find that the data looks pretty nice. Due to the page limit, we do not show the plots. Therefore, we decide not to delete the outliers in Figure 0.

# Part 1: Exploring the data

In this part, we will answer the questions given in the guidance.

## Question 1

**Description**

In 2019, what region of the world has the highest rate of alcohol-related deaths among men aged 40-44?

**Method**

We extract the data of the alcohol-related death rate among men aged 40-44 and sort the location according to the death rate in the descending order.

```{r}
highest_alcohol_deaths = substance_use %>%
  filter(measure == "Deaths", year == 2019, age == "40 to 44",
         sex == "Male", cause == "Alcohol use disorders") %>%
  select(location, val) %>%
  arrange(desc(val)) %>%
  top_n(1, val)

kable(highest_alcohol_deaths)
```

**Conclusion**

In 2019, Europe & Central Asia has the highest rate of alcohol-related deaths among men aged 40-44.

## Question 2

**Description**

Looking at the prevalence of alcohol-related disease in the East Asia and Pacific region, how has this changed over time and in the different age groups? Is there a difference between men and women?

**Method**

First, we extract the data.

```{r}
eap_alcohol_data = substance_use %>%
  filter(measure == "Prevalence", cause == "Alcohol use disorders",
         location == "East Asia & Pacific - WB") %>%
  select(year, age, sex, val)
```

Next, we visualize the data in two ways, faceting the figures by age groups and by sex groups.

```{r}
eap_alcohol_data1 = eap_alcohol_data
eap_alcohol_data1$age = paste0(eap_alcohol_data1$age, " years old")
ggplot(eap_alcohol_data1,
       aes(x = year, y = val, color = sex,
           group = interaction(sex, age))) +
  geom_line() +
  facet_wrap(~age, scales = 'free_y') +
  labs(x = "Year", y = "Prevalence (%)", color = "Sex") +
  theme_bw() +
  theme(legend.position = "right",
        plot.title = element_text(hjust = 0.5))
```

Figure 2.1: The trends of the alcohol-related disease prevalence in east Asia and Pacific. This figure is faceted by age groups, better for visually comparing the prevalence between the male and the female.

```{r}
ggplot(eap_alcohol_data,
       aes(x = year, y = val, color = age,
           group = interaction(sex, age))) +
  geom_line() +
  geom_point(size = 0.6) +
  facet_wrap(~sex, scales = 'free_y') +
  labs(x = "Year", y = "Prevalence (%)", color = "Age (years old)") +
  theme_bw() +
  theme(legend.title = element_text(hjust = 0.5),
        plot.title = element_text(hjust = 0.5))
```

Figure 2.2: The trends of the alcohol-related disease prevalence in east Asia and Pacific. This figure is faceted by sex groups, better for visually comparing the prevalence between different age groups.

From Figure 2.1 and Figure 2.2, you can see how the prevalence of the alcohol-related disease in the East Asia and Pacific region changes over time in different age groups. Also, you can identify different patterns of the prevalence between male and female.

In the following, we will validate the difference between male and female in a more statistical way.

First, we formulate the hypotheses.

-   The null hypothesis (H0): There is no difference in the prevalence of different age groups between male and female.
-   The alternative hypothesis (HA): There is difference in the prevalence of different age groups between male and female.

The data for male and female in the same year paired, and we want to test the differences in prevalence between male and female, so we will use the paired statistical test in the following. First, We test the normality of the differences in prevalence.

-   If the differences are normally distributed, we will perform the parametric Student's t-test to test the hypotheses.
-   If the differences are not normally distributed, we will perform the non-parametric Wilcoxon rank sum test to test the hypotheses.

```{r}
eap_alcohol_data2 = eap_alcohol_data %>%
  group_by(age) %>%
  group_split() %>%
  map(~spread(., key = "sex", value = "val")) %>%
  map(~{
    shapiro_p = shapiro.test(.$Male - .$Female)$p.value
    if(shapiro_p < 0.05){
      test_type = "Wilcoxon rank sum test"
      p_value = wilcox.test(.$Male, .$Female, paired = T)$p.value
    }
    else{
      test_type = "Student's t-test"
      p_value = t.test(.$Male, .$Female, paired = T)$p.value
    }
    tibble(shapiro_p = shapiro_p, test_type = test_type, p_value = p_value)
  }) %>%
  bind_rows() %>%
  mutate(age = unique(eap_alcohol_data$age), .before = shapiro_p)
kable(eap_alcohol_data2, digit = 12)
```

**Conclusion**

Of all the age groups, the p values from the statistical test are all less than $10^{-8}$. Therefore, we can reject the null hypothesis (H0). There is sufficient evidence to conclude that there is significant difference in the prevalence between the male and the female within each age group.

## Question 3

**Description**

In the United States, there is talk of an “Opioid epidemic”. Part of the problem is that since the late 1990s, doctors have increasingly been prescribing pain killers which can be highly addictive. Looking at the data from the United States, can you confirm an increase in the prevalence of diseases related to opioid use? What age group is the most affected?

**Method**

There is no data named "United States" in the locations divided by the world bank, so we extract the data from "North America" to represent the data from the United States.

According to the question, we will separate the data from 1990 to 2019 into two parts "before the start of increasing prescription of the opioid" (1990-1997) and "after the start of increasing prescription of the opioid" (1998-2019). We will abbreviate the two periods to the "before" period and the "after" period in the following.

```{r}
opioid_use_na = substance_use %>%
  filter(measure == "Prevalence",
         location == "North America",
         cause == "Opioid use disorders") %>%
  select(age, year, sex, val)
```

Next, we visualize the data in two ways, faceting the figures by age groups and by sex groups.

```{r}
opioid_use_na1 = opioid_use_na
opioid_use_na1$age = paste0(opioid_use_na1$age, " years old")
ggplot(opioid_use_na1,
       aes(x = year, y = val, color = sex,
           group = interaction(sex, age))) +
  geom_line() +
  geom_vline(xintercept = 1998, color = "darkgray") +
  facet_wrap(~age, scales = 'free_y') +
  labs(x = "Year", y = "Prevalence (%)", color = "Sex") +
  theme_bw() +
  theme(legend.position = "right",
        plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45))
```

Figure 3.1: The trends of the opioid use related disease prevalence in North America. The figure is faceted by age groups, better for visually comparing the prevalence between male and female.

```{r}
ggplot(opioid_use_na,
       aes(x = year, y = val, color = age,
           group = interaction(sex, age))) +
  geom_line() +
  geom_point(size = 0.7) +
  geom_vline(xintercept = 1998, color = "darkgrey") +
  facet_wrap(~sex, scales = 'free_y') +
  labs(x = "Year", y = "Prevalence (%)", color = "Age (years old)") +
  theme_bw() +
  theme(legend.title = element_text(hjust = 0.5),
        plot.title = element_text(hjust = 0.5))
```

Figure 3.2: The trends of the opioid use related disease prevalence in North America. The figure is faceted by sex groups, better for visually comparing the prevalence between different age groups.

From Figure 3.1 and Figure 3.2, we can identify that there is an overall increase pattern within each age and sex group. Since the prevalence of the "before" and "after" period are both increasing, and we want to confirm whether the increasing prescription the opioid has an effect on the prevalence, we should compare the speed (slope) of the increase in the prevalence rather than the prevalence itself. If the prevalence of the "after" period is increasing faster than the prevalence of the "before" period, the increasing prescription of the opioid may have an effect.

To test whether the slope of the prevalence of the "before" period is different from the slope of the prevalence of the "after" period, we formulate the hypotheses.

-   The null hypothesis (H0): There is no difference in the slope of the prevalence between the "before" period and the "after" period.
-   The alternative hypothesis (HA): There is a difference in the slope of the prevalence between the "before" period and the "after" period.

Because the sampling is random, we will use the two-tailed Wilcoxon rank sum test to test the hypotheses.

```{r}
opioid_use_na2 = opioid_use_na %>%
  group_by(sex, age) %>%
  group_split() %>%
  map(~{
    before = diff(.$val[.$year < 1998])
    after = diff(.$val[.$year >= 1998])
    # calculate the slope
    tibble(
      sex = .$sex[1],
      age = .$age[1],
      p_value = wilcox.test(after, before)$p.value,
      effect_size = median(after) - median(before)
    )
  }) %>%
  bind_rows() %>%
  group_by(sex) %>%
  group_split()
kable(opioid_use_na2)
```

Table 3.1: The age (age), the sex (sex), the p value (p_value) and the effect size (effect_size) of the the Wilcoxon rank sum test for each group.

From Table 3.1, within each age and sex group, the p value of the Wilcoxon rank sum test is larger than 0.05, we cannot reject the null hypothesis (H0).

However, if we do not consider the sex as a covariate, we will take the average of the prevalence of male and female to represent the prevalence of the whole population within each age group, as the ratio of the male population and the female population within each age group of the divided locations is approximately 1:1.

```{r}
opioid_use_na3 = opioid_use_na %>%
  group_by(year, age) %>%
  summarize(val = mean(val)) %>%
  group_by(age) %>%
  group_split() %>%
  map(~{
    before = diff(.$val[.$year < 1998])
    after = diff(.$val[.$year >= 1998])
    # calculate the slope
    tibble(
      age = .$age[1],
      p_value = wilcox.test(after, before)$p.value,
      effect_size = median(after) - median(before)
    )
  }) %>%
  bind_rows()
kable(opioid_use_na3)
```

Table 3.2: The age (age), the p value (p_value) and the effect size (effect_size) of the the Wilcoxon rank sum test for each group.

From Table 3.2, when we do not take the sex as a covariate, we can identify that except 25-29 age group, the p-values for the other age groups are all smaller than 0.05, so we can reject the null hypothesis (H0). Therefore, except 25-29 age group, there is sufficient evidence to conclude that there is significant difference in the slope of the prevalence between the "before" period and the "after" period. Additionally, as the effect size (after - before) are all larger than 0, so we can confirm that the increasing prescription of the opioid can accelerate the increase in the prevalence of the opioid use disease except 25-29 age group.

From Figure 3.1 and Figure 3.2, we can assume that there is a linear relationship between the year and the prevalence. We use the linear regression model to fit the data, and we test the normality and the homoscedastic of the residuals. To find which age group is the most affected, we compare the difference in the slope of the "before" and "after" period except 25-29 age group.

As the "Multiple R-squared" represents the proportion of the variability of the dependent variable explained by the model, we set the threshold of the "Multiple R-squared" as 0.7 (recommended in the ADS2 Lecture 2.6 Slide Page 29).

```{r}
opioid_use_na4 = opioid_use_na %>%
  filter(age != "25 to 29") %>%
  group_by(year, age) %>%
  summarize(val = mean(val)) %>%
  group_by(age) %>%
  group_split() %>%
  map(~{
    before_data = filter(., year < 1998)
    after_data = filter(., year >= 1998)
    before = summary(lm(val ~ year, before_data))
    after = summary(lm(val ~ year, after_data))
    tibble(age = .$age[1],
           before_r_squ = before$r.squared,
           before_slope = before$coefficients[2],
           after_r_squ = after$r.squared,
           after_slope = after$coefficients[2],
           slope_diff = after_slope - before_slope
           )
  }) %>%
  bind_rows() %>%
  arrange(desc(slope_diff))
kable(opioid_use_na4)
```

Table 3.3: The age (age), the "Multiple r-squared" of the "before" period (before_r_squ), the slope of the prevalence in the "before" period (before_slope), the "Multiple r-squared" of the "after" period (after_r_squ), the slope of the prevalence in the "after" period (after_slope), and the difference in the slope between the "before" and "after" period (slope_diff).

From Table 3.3, because the "Multiple r-squared" values are all greater than 0.7 except 65-69 age group (the "before" period), we can compare the difference in the slope between the "before" and "after" period except 65-69 age group. The 30-34 age group has the largest difference in the slope of the prevalence of the opioid use disease, so the 30-34 age group is the most affected.

You can use the following code to further verify the normality and homoscedastic of the residuals of this model. Due to the page limit, we do not show the results of the code.

```{r, eval = FALSE}
plot(model, c(1, 2))
```

**Conclusion**

If we take the sex as a covariate, we cannot confirm the increasing prescription of the opioid can accelerate the increase in the prevalence of the opioid use disease in each age group. If we do not take the sex as a covariate, we can confirm that the increasing prescription of the opioid can accelerate the increase in the prevalence of the opioid use disease except 25-29 age group, and the 30-34 age group is the most affected.

# Part 2: Ask your own question

First, in Figure 2.1, we identify that in east Asia and Pacific, for all age groups, the alcohol-related disease prevalence of the male is higher than that of the female during the years. We are interested in whether in different locations in the world, the alcohol-related death rate of the male is higher than the female. Then, we do some visualization.

```{r}
data1 = substance_use %>%
  filter(measure == "Deaths",
         cause == "Alcohol use disorders") %>%
  select(location, age, year, sex, val)

data1$location = gsub(data1$location, pattern = " - WB", replacement = "")

ggplot(data1, aes(x = location, y = val, fill = location)) +
  geom_boxplot(outlier.size = 0.1) +
  labs(x = "Age", y = "Difference in death rate (%)") +
   theme_bw() +
  facet_wrap(~sex) +
  theme(axis.text.x = element_blank())
```

Figure 4.1 The the alcohol-related death rate of the male and the female in different locations in the world.

From Figure 4.1, we find that the alcohol-related death rates in Sub-Saharan Africa and Middle East & North Africa are much lower than the other locations for both male and female, so there is little significance to compare the difference in death rate in these two locations. we regard them as outliers and drop them out. Here we give a possible explanation: both Sub-Saharan Africa and Middle East & North Africa are mostly “underdeveloped areas”. People in these places have few chances to have a heavy alcohol exposure.

Next, we compare the difference in the death rate of the male and female in the selected locations.

```{r}
locations = c("East Asia & Pacific", "North America", "South Asia", 
"Latin America & Caribbean", "Europe & Central Asia")

data2 = data1 %>%
  filter(location %in% locations) %>%
  group_by(location, age, year) %>%
  spread(key = "sex", val = "val") %>%
  summarize(diff = Male - Female) %>%
  bind_rows()

ggplot(data2, aes(x = age, y = diff, fill = age)) +
  geom_boxplot(outlier.size = 0.1) +
  facet_wrap(~location, scale = "free_y") +
  labs(x = "Age", y = "Difference in death rate (%)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position = "none")
```

Figure 4.2 The difference in the alcohol-related death rate between male and female in different locations.

From Figure 4.2, we find that in different locations, the differences in the prevalence between the male and the female in different age groups have different distributions. We are interested in whether the location and the age have any interaction when they affect the difference in the death rate.

Therefore, our question is

**Do the location and the age have any interaction when they affect the difference in the alcohol-related death rate between the male and the female?**

First, we formulate the hypotheses.

-   The null hypothesis (H0): There is no interaction between the location and the age.
-   The alternative hypothesis (HA): There is an interaction between the location and the age.

First, we build a model. As the sampling is random, we test the normality and homoscedastic of the residuals.

```{r}
model2 = lm(diff ~ location * age, data = data2)
par(mfrow = c(1, 2))
plot(model2, c(1, 2),cex=0.4)
par(mfrow = c(1, 1))
summary(model2)$r.squared
anova(model2)
```

From the Q-Q plot, we can find that the residuals are normally distributed. From the Residuals vs Fitted plot, we can find that the residuals are homoscedastic.

The "Multiple R-squared" value represents the proportion of the variability of the dependent variable explained by the model. The "Multiple R-squared" value is 0.9635071, so the model explains the data quite well. The p-value of ANOVA is smaller than 2.2e-16, so we can reject the null hypothesis (H0). There is sufficient evidence to conclude that there is a significant interaction between the location and the age.

**Conclusion**

There is a significant interaction between the location and the age when they affect the difference in the alcohol-related death rate between the male and the female.
