---
title: 'Review of coding challenge for Semester 1'
author: '2118'
date: "2024-01-10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tinytex)
library(rmarkdown)
library(knitr)
library(tidyverse)
library(cowplot)
setwd("E:\\A大学\\大二下\\ADS2\\W17 Review for Semester 1")
```

# 1. Benefits of swimming for long-distance runners

```{r}
swim = read.table("swimming.txt", sep = '\t', header = T)
# head(swimming)
# summary(swim)
# str(swim)
```

## 1.1 Tidy the data and decide on suitable statistical test.

### How are the before and after time distributed?

```{r}
before_time = swim$before_minutes * 60 + swim$before_seconds
before_time = as.integer(before_time)
after_time = swim$after_minutes * 60 + swim$after_seconds
after_time = as.integer(after_time)
hist(before_time - after_time, breaks = 20)
shapiro.test(before_time - after_time)
swim2 = data.frame(swim$names, before_time, after_time)
# head(swim2)
```

- The improvement time (the difference between
the after_time and the before_time) are all normally distributed.
- However, we cannot decide whether the improvement time is normally distributed
  through the histogram.
- It is better to use the shapiro test to test the normality.
- In the shapiro test, p > 0.05, so the improvement is normally distributed,
  so we can use the t test.
- Since every person in the data has a before_time and an after_time, the 2
values are paired.
- As we want to test the improvement time, We decided to use the paired 2-sample t-test. (NOTE: To get full mark, you should emphasize that we test the improvement time.)

## 1.2 The null and alternative hypotheses

- The null hypothesis (H0): The time used for the half-marathon after
the swimming training is no shorter than that before the swimming training.
- The alternative hypothesis (HA) : The time used for the half-marathon after
the swimming training is shorter than that before the swimming training.

## 1.3 Is there a statistically significant improvement on runners’ times after swimming?

```{r}
t.test(after_time, before_time, paired = T, alternative = "less")
t.test(after_time - before_time, mu = 0)
```

- NOTICE: t.test(x, y, alternative = 'greater') means the HA is y>x
- p > 0.05
- We cannot reject the null hypothesis.
- There is insufficient evidence to conclude that the time used for the half-marathon after the swimming training is shorter than that before the
swimming training.
- Therefore, there is not a statistically significant improvement on runners’ times after swimming.



# 2. Number of emergency room admissions

## 2.1 Import the dataset and plot the data in a useful way

```{r}
hosp = read.csv("hospital_admissions.csv")
# head(hosp)
# str(hosp)
hosp$week = as.factor(hosp$week)
hosp$weekday = as.factor(hosp$weekday)

subset(hosp, hosp$week == 1)

hosp1 = aggregate(hosp$patients_per_hour,
                  by = list(hosp$week, hosp$weekday),
                  FUN = sum)
names(hosp1)[1] = "week"
names(hosp1)[2] = "weekday"
names(hosp1)[3] = "patients"
# str(hosp1)
# summary(hosp1)
g2.1 = ggplot(data = hosp1[hosp1$weekday == "Monday",],
            mapping = aes(x = week, y = patients)
            )
g2.1 = g2.1 + geom_bar(stat="identity", fill = "orange")
g2.1
```

\begin{center}
Figure 2.1: Patients on Monday during the year.
\end{center}

```{r}
g2.2 = ggplot(data = hosp1[hosp1$weekday == "Sunday",],
            mapping = aes(x = week, y = patients)
            )
g2.2 = g2.2 + geom_bar(stat = "identity", fill = "purple")
g2.2
```

\begin{center}
Figure 2.2: Patients on Sunday during the year.
\end{center}

- We can change the x axis into patients_per_day and the y axis into the count.
- In this way, we remove the information of week, because we think the data are not paired, which means the patient number on Sunday and Monday are not correlated.
- This process is like to put the original patient number data into buckets to remove the week information.

```{r}
#hosp1
g2.3 = ggplot(hosp1, mapping = aes(x = patients, fill = weekday))
g2.3 = g2.3 + geom_histogram(position = "identity", alpha = 0.2, bins = 10)
g2.3
```

- You can see that the patients on Sundays are normally distributed. Also, the patients on Mondays are normally distributed.

```{r}
boxplot(hosp1$patients ~ hosp1$weekday, notch = T, ylab = "patients per day")

```

## Is there a difference in patient admission rates between Mondays and Sundays?

- We first form the null (H0) and alternative (HA) hypothesis for this question.
- H0: The patient admission rate on Mondays is no different from that on Sundays.
- HA: The patient admission rate on Mondays is different from that on Sundays.

```{r}
monsums = hosp1[hosp1$weekday == "Monday", ]$patients
sunsums = hosp1[hosp1$weekday == "Sunday", ]$patients
shapiro.test(monsums)
shapiro.test(sunsums)
```
- The patients on both Monday and Sunday during the year are not normally distribution, so the data do not meet the condition to use the t-test.

```{r}
wilcox.test(hosp1[hosp1$weekday == "Monday", ]$patients, 
            hosp1[hosp1$weekday == "Sunday", ]$patients,
            paired = F) 
```

- p-value < 0.05
- We reject H0.
- There is sufficient evidence to conclude that
the patient admission rate on Mondays is different from that on Sundays.
- Therefore, there is a significant difference in patient admission rates between Mondays and Sundays.

- Here provide another solution: simulation.
```{r}
monsums = hosp1[hosp1$weekday == "Monday", ]$patients
sunsums = hosp1[hosp1$weekday == "Sunday", ]$patients

real_median = median(monsums) - median(sunsums)

random_sums = c(sunsums, monsums)

count = 0

reptimes = 10000

for(i in 1:reptimes){
  random_index = sample(seq(1,104), 52, replace = F)

  random_mon = random_sums[random_index]
  random_sun = random_sums[-random_index]
  
  random_median = median(random_mon) - median(random_sun)
  
  if(random_median >= real_median)
    count = count + 1
}

p_value = 1.0*count/reptimes
p_value
```
- p-value < 0.05, so we reject H0.
- NOTE: p-value means the probability that the situation is the same or more extreme (refers to extreme like HA) if H0 is true.

```{r}
#monsums
#sort(monsum)
plot(sort(monsums), 1:length(monsums)/length(monsums), type = 'l',
     ylab = "cumulative proportion", xlab = "patients per day")
lines(sort(sunsums), 1:length(sunsums)/length(sunsums), col = "red")
legend("bottomright", legend = c("Sunday", "Monday"), fill = c("red", "black"),
       bty = "n")
# bty = box type, default = "o", "n" means none.
```

## 2.3 Based on your findings, what advice would you give Dr. Horsey?

- We should arrange more staff on Mondays than on Sundays.


# 3. Spinal cord injury and novel biomaterials

## 3.1 Import, arrange the data (merge both pieces of data and make the data possible to analyse), and makeit suitable for analysis.

```{r}
data1 = read.csv("SCI_before.csv")
data2 = read.csv("SCI_after.csv")
# head(data1)
# head(data2)
data1$patient_ID = as.factor(data1$patient_ID)
# levels(data1$patient_ID)
data2$patient_ID = as.factor(data2$patient_ID)
# summary(data1)
# summary(data2)
# library(dplyr)
# levels(data2$patient_ID)
data1 = arrange(data1, data1$patient_ID)
data2 = arrange(data2, data2$patient_ID)
# data1$patient_ID == data2$patient_ID
#data1
#data2
#match(data1$patient_ID, data2$patient_ID)
data = cbind(data1, data2$AIS_after)
names(data)[3] = "AIS_after"
# head(data)
```

## Any NA?

```{r}
anyNA(data)
```

- No NA.

## Any duplicated?
```{r}
idx1 = which(duplicated(data))
idx2 = which(duplicated(data, fromLast = T))
idx1
# data[c(idx1, idx2),  ]
data = data[-idx1, ]
```

### Data type?
```{r}
# summary(data)
data$AIS_before = as.factor(data$AIS_before)
data$AIS_after = as.factor(data$AIS_after)
```

## Documentation: Remove 8 duplicated rows.

## 3.2 Check your data carefully. Identify features of the data and discuss your conclusions. Make illustrative plots.

```{r}
g3.1 = ggplot(data = data,
            mapping = aes(x = AIS_before))
g3.1 = g3.1 + geom_bar(fill = "orange")
g3.1
```

\begin{center}
Figure 3.1: ALS level distribution before treatment.
\end{center}

```{r}
g3.2 = ggplot(data = data,
            mapping = aes(x = AIS_after))
g3.2 = g3.2 + geom_bar(fill = "purple")
g3.2
```

\begin{center}
Figure 3.1: ALS level distribution after treatment.
\end{center}


## 3.3 Formulate the correct statistical hypothesis to compare the groups, choose the appropriate statistical test

- We first form the null (H0) and alternative (HA) hypothesis for this question.
- H0: The AIS score after treatment is no better than that before treatment.
- HA: The AIS score after treatment is better than that before treatment.
- Because the sample size is too small, we cannot decide whether it is normally distributed.
- Therefore, We use paired Wilcoxon test.
- We convert AIS score A,B,C,D,E into 5, 4, 3, 2 and 1.

```{r}
a = c()
for (i in 1:nrow(data)) {
  x = data[i, "AIS_before"]
  if (x == "A")
    t = 5
  if (x == "B")
    t = 4
  if (x == "C")
    t = 3
  if (x == "D")
    t = 2
  if (x == "E")
    t = 1
  a = c(a, t)
}
#print(a)
b = c()
for (i in 1:nrow(data)) {
  x = data[i, "AIS_after"]
  if (x == "A")
    t = 5
  if (x == "B")
    t = 4
  if (x == "C")
    t = 3
  if (x == "D")
    t = 2
  if (x == "E")
    t = 1
  b = c(b, t)
}
wilcox.test(a, b, alternative = 'less', paired = T)
```
- p-value < 0.05
- We reject H0.
- There is sufficient evidence to conclude that the AIS score after treatment is better than that before treatment.

```{r}
data_to_plot = cbind(rep(1, nrow(data)),a,rep(2, nrow(data)),b)
data_to_plot = as.data.frame(data_to_plot)
names(data_to_plot) = c("AIS_before", "x0", "AIS_after", "x1")
# str(data_to_plot)
data_to_plot[, c(1, 3)] = apply(data_to_plot[, c(1, 3)], MARGIN = c(1, 2), FUN = jitter, factor = 0.5)
# MARGIN = for a matrix 1 indicates rows, 2 indicates columns, c(1, 2) indicates rows and columns
data_to_plot
plot(x = data_to_plot[ , 1], y = data_to_plot[ , 2],
     xlim = c(0.5, 2.5), ylim = c(2.5, 5.5),
     axes = F, xlab = "", ylab = "Disability score")
# add the point of AIS before AND hide the axis
points(x = data_to_plot[ , 3], y = data_to_plot[ , 4])
# add the point of AIS after
segments(x0 = data_to_plot[ , 1],
         y0 = data_to_plot[ , 2],
         x1 = data_to_plot[ , 3],
         y1 = data_to_plot[ , 4])
axis(2) # add y axis
axis(1, at = 1:2, labels = c("AIS_before", "AIS_after"))
# only draw the x axis between (1, 2)
box(bty = "l")
# bty = L means the frame looks like the letter 'L'
# add a box around the plot, the aim is to complete the x axis
```

```{r}
row.names(data) = 1:nrow(data)
data$AIS_before = a
data$AIS_after = b
data = gather(data, key = "measurement", value = "AIS", AIS_before, AIS_after)
data$patient_ID = as.factor(data$patient_ID)
model = lm(AIS ~ measurement + patient_ID, data) #
# summary(model)
# shapiro.test(resid(model))
# use q-q plot (quantile-quantile plot) to validate normality.
qqdata = resid(model)
# summary(qqdata)
qqnorm(qqdata)
qqline(qqdata, col = "red", lwd = 2) # add a reference line
xlab("Theoretical Quantiles")
ylab("Ordered Values")
```
- Normality check shows that the residuals from a linear regression of the data     are not normally distributed.
- Because most points are not tightly cluster around the reference line.

## 3.4 Discuss the results you got.

- The treatment has significant improvement effect on the spinal cord injury.
- The effect size.

```{r}
effect_size = abs(mean(b)-mean(a))
effect_size
```
- We should consider more factors that influence the AIS score.
- We should consider to use a quantitative score to evaluate the health condition
instead of AIS score.